---
title: "Other Datasets"
author: "Len Huang"
date: "5/18/2021"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(tidyr)
library(tidyverse)
setwd("~/Coding/School/73265/73265-Final-Project")
```

# Combining Datasets

## Preparing Our Data

Dataset from here: https://www.kaggle.com/kumarajarshi/life-expectancy-who. We 
keep the columns: country, life expectancy, and gdp per capita.

```{r}
life <- read_csv("life.csv")
life <- life[c(1,4,17)]
life <- drop_na(life)
```

We group the data together by averaging out each country's information. These
will be aggregated over the years from 2000 - 2015

```{r}
life <- aggregate(life[, 2:3], list(life$Country), mean)
names(life) <- c("Country", "LifeExpectancy", "GDPperCapita")
```

Do the same for our happiness dataset:

```{r}
data2015 <- read_csv("archive/2015.csv")
data2019 <- read_csv("archive/2019.csv")
names(data2015) <- c("Country", "Region", "Rank", "Happiness", "SE",
                     "Economy", "Social", "Health", "Freedom", "Trust",
                     "Generosity", "Dystopia")
names(data2019) <- c("Rank", "Country", "Happiness", "Economy", "Social",
                     "Health", "Freedom", "Generosity", "Trust")
data2015 <- data2015 %>% select("Country", "Happiness", "Economy", "Health")
data2019 <- data2019 %>% select("Country", "Happiness", "Economy", "Health")

databoth <- rbind(data2015, data2019)

databoth <- aggregate(databoth[, 2:4], list(databoth$Country), mean)

names(databoth) <- c("Country", "Happiness", "EconomyContr", "HealthContr")
```

Now let's try to combine our dataset from happiness with our dataset of life
expectancy and gdp per capita.

```{r}
data <- merge(life, databoth, by="Country")
data
```

## Exploring Our Data

Let's make this longer and more graphable 

```{r}
data %>%
  select(-c("Country")) %>%
  arrange(desc(Happiness)) %>%
  pivot_longer(cols = c("LifeExpectancy", "GDPperCapita", "Happiness", 
                        "EconomyContr", "HealthContr"),
               names_to = "Category",
               values_to = "Score")
```

Just kidding may not need that, but let's try happiness vs contribution and 
happiness vs actual for both categories (GDP / Health).

### Health

```{r}
data %>% ggplot(aes(x=LifeExpectancy, y=Happiness)) + 
  geom_point() + geom_smooth() + 
  theme_bw() +
  ggtitle("Happiness Score vs Life Expectancy") 

data %>% ggplot(aes(x=HealthContr, y=Happiness)) + 
  geom_point() + geom_smooth() + 
  theme_bw() +
  ggtitle("Happiness Score vs WHO Contribution Score for Health") 
```

### GDP

```{r}
data %>% ggplot(aes(x=GDPperCapita, y=Happiness)) + 
  geom_point() + geom_smooth() + 
  theme_bw() +
  ggtitle("Happiness Score vs GDP Per Capita") 
data %>% ggplot(aes(x=EconomyContr, y=Happiness)) + 
  geom_point() + geom_smooth() + 
  theme_bw() +
  ggtitle("Happiness Score vs WHO Contribution Score for Economy") 
```